Quantifying Lewontin's Paradox Suggests Natural Selection is Unlikely to Explain the Narrow Range of Diversity Among Species

### 2021-06-04 manuscript/elife/elife-template.tex
even more elife formatting bullshit

   \figsupp[The relationship between body mass and population density found by
   \cite{Damuth1987-sg}, which is used to predict population densities.]{The
     source of this data is appendix table of \cite{Damuth1987-sg}; the color
     indicates Damuth's original group labels. The dashed line was estimated
     using a lognormal regression model in Stan. References to each measurement
   are available in \cite{Damuth1987-sg}.}{\includegraphics[]{figures/damuth.pdf}}\label{figsupp:sf1}

@@

@@

A major impediment in quantifying Lewontin's Paradox has been estimating census
population sizes across many taxa, especially for extremely abundant,
cosmopolitan species that define the upper limit of ranges. Previous work has
surveyed the literature for census size estimates
\citep{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range, body size, or
qualitative categories as proxies for census size
\citep{Corbett-Detig2015-gt,Leffler2012-zj}. To quantify the relationship
between genomic estimates of diversity and census population sizes, I first
approximate census population sizes for 172 metazoan taxa (\FIG{figure-nc}). My
approach predicts population densities from body sizes using a
previously-observed linear relationship that holds across metazoans (see
\FIGSUPP[figure-nc]{sf1}; \cite{Damuth1981-st,Damuth1987-sg}). Then, from
geographic occurrence data, I estimate range sizes.  Finally, I estimate
population size as the product of these predicted densities and range estimates
(see Methods and Materials: \nameref{sec:methods-popsize}). Note that the relationship between
population density and body size is driven by energy budgets, and thus reflects
macroecological equilibria \citep{Damuth1987-sg}. Consequently, population
sizes are underestimated for taxa like humans and their domesticated species,
and overestimated for species with anthropogenically reduced densities or
fragmented ranges. For example, the population size of \emph{Lynx lynx} is
likely around 50,000 \citep{Iucn2020-ap} which is around two orders of
magnitude smaller than my estimate. Additionally, the range size estimates do
not consider whether an area has unsuitable habitat, and thus may be
overestimated for species with particular niches or patchy habitats. While
these methods to estimate census size are crude and approximate, they can be
efficiently calculated for numerous taxa and are sufficient to estimate the
scale of Lewontin's Paradox (see \nameref{supp:popsize-valid} for more on
validation based on biomass and other approaches).

@@

\begin{align} 
  \pi_\text{BGS+HH} \approx \frac{\theta}{\nicefrac{1}{B(U, L)} + 2N_c S(\gamma, L, J)} 
  \label{eqn:div} 
\end{align}
%
where $\theta = 4N_c\mu$, $B(U,L)$ is the effect of background selection, and
$S(\gamma, L, J)$ is the rate of coalescence caused by sweeps (c.f.
\cite{Elyashiv2016-vt}, equation 1, \cite{Coop2012-cd} equation 20). Under
background selection models with recombination, the reduction is $B(U,L) =
\exp(\nicefrac{-U}{L})$ where $U$ is the per diploid genome per generation
deleterious mutation rate, and $L$ is the recombination map length
\citep{Hudson1994-oh,Hudson1995-xc,Nordborg1996-nq}. This BGS model is
similar to models of effective population size under polygenic fitness
variation, and can account for other modes of linked selection
(\cite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}, see Appendix Section
\ref{app:bgs}). The coalescent rate due to sweeps is $S(\gamma, L, J) =
\nicefrac{\gamma}{L} J$, where $\gamma$ is the number of adaptive substitutions
per generation, and $J$ is the probability a lineage is trapped by sweeps as
they occur across the genome (c.f. $J_{2,2}$ in equation 15 of
\cite{Coop2012-cd}).

@@

  \caption{(A) The observed relationship between recombination map length ($L$)
    and census size ($N_c$) across 136 species with complete data and known
    phylogeny. Triangle points indicate six social taxa excluded from the model
    fitting since these have adaptively higher recombination map lengths
    \citep{Wilfert2007-dx}. The dark gray line is the estimated
    relationship under a phylogenetic mixed-effects model, and the gray
    interval is the 95\% posterior average. (B) Points indicate the observed
    $\pi$--$N_c$ relationship across taxa shown in Figure \ref{fig:figure-1},
    and the blue ribbon is the range of predicted diversity were $N_e = N_c$
    for $\mu = 10^{-8}$--$10^{-9}$,  and after accounting for the expected
  reduction in diversity due to background selection and recurrent hitchhiking
under \emph{Drosophila melanogaster} parameters. In both plots, point color
indicates phylum.}\label{fig:linksel}

@@

Previous work has found that the impact of linked selection increases with
$N_c$ (\cite{Corbett-Detig2015-gt}; see also Supplementary Figure
\ref{figsupp:corbett_detig}A), and it is often thought that this is driven by
higher rates of adaptive substitutions in larger populations
\citep{Ohta1992-yi}, despite equivocal evidence \citep{Galtier2016-dq}.
However, there is another mechanism by which species with larger population
sizes might experience a greater impact of linked selection: recombinational
map length, $L$, is known to correlate with body mass \citep{Burt1987-tq} and
thus varies inversely with population size. As this is a critical parameter
that determines the genome-wide impact of both hitchhiking and background
selection, I examine the relationship between recombination map length ($L$)
and census population size ($N_c$) across taxa, using available estimates of
map lengths across species \citep{Stapley2017-fs,Corbett-Detig2015-gt}. I find
a significant non-linear relationship using phylogenetic mixed-effects models
(Figure \ref{fig:figure-3}A; see Methods and Materials:
\nameref{sec:methods-pcm}).  There is also a correlation between map length and
genome size (Supplementary Figure \ref{supfig:genome_size_maplength}) and
genome size and population size (Supplementary Figure
\ref{supfig:genome_size_popsize}). These findings are consistent with both the
hypothesis that non-adaptive processes increase genome size in small-$N_e$
species \citep{Lynch2003-bn} which in turn could increase map lengths, as well
as the hypothesis that map lengths are adaptively longer to more efficiently
select against deleterious alleles \citep{Roze2021-ya}. Overall, the negative
relationship between map length and census size indicates linked selection is
expected to be stronger in short map length, high-$N_c$ species.

Then, I predict the expected diversity ($\pi_{BGS+HH}$) under background
selection and hitchhiking, were $N_e= N_c$, and assuming all species had the
rate of sweeps and strength of BGS as \emph{D. melanogaster}. Since neutral
mutation rates $\mu$ are unknown and vary across species, I calculate the range
of predicted $\pi_{BGS+HH}$ estimates for $\mu = 10^{-8}$--$10^{-9}$ (using the
four-alleles model, \cite{Tajima1996-rb}), and compare this to the observed
relationship between $\pi$ and $N_c$ in Figured \ref{fig:figure-3}B.  Under
these parameters, linked selection begins to appreciably depress diversity
around $N_c \approx 10^9$, since $S \approx 10^{-8}$--$10^{-9}$ and linked
selection dominates drift when $S > \nicefrac{1}{2N}$.  Overall, this reveals
two problems for the hypothesis that linked selection could solve Lewontin's
Paradox. First, low to mid-$N_c$ species (census sizes between
$10^{4}$--$10^{10}$) have sufficiently long map lengths that their diversity
levels are only moderately reduced by linked selection, leading to a wide gap
between predicted and observed diversity levels. For this not to be the case,
the parameters that determine the strength of background selection and
recurrent hitchhiking would need to be \emph{higher} among these species than
in \emph{Drosophila melanogaster}. This would require that the rate of adaptive
mutations or the deleterious mutation rate be orders of magnitude higher for
species within this range than in \emph{Drosophila}, which is incompatible with
the rate of adaptive substitutions across species \citep{Galtier2016-dq}
and mutation rates \citep{Lynch2010-ki}. Furthermore, linked selection has
been quantified in humans, which fall in this census size range, and has been
found to be relatively weak
\citep{McVicker2009-ax,Hernandez2011-gs,Hellmann2008-ic,Cai2009-by,Boyko2008-tj}.
Second, while hitchhiking and BGS can reduce predicted diversity levels for
high-$N_c$ species ($N_c > 10^{10}$) to observed levels, this would imply
available estimates of $\pi_0$ are underestimated by several orders of
magnitude in \emph{Drosophila} (Supplementary Figured
\ref{figsupp:corbett_detig}B). The high reductions in $\pi$ predicted here
(compared to those of \cite{Elyashiv2016-vt}) are a result of using $N_c$,
rather than $N_e = \nicefrac{\pi_0}{4 \mu}$ in the denominator of Equation
\eqref{eqn:div}, which leads to a very high rate of sweeps in the population.
I do not consider selective interference, though the saturation of adaptive
substitutions per Morgan would only act to limit the reduction in diversity
\citep{Weissman2012-aa}, and thus these results are conservative.
Overall, while linked selection could decouple diversity from census size for
high-$N_c$ species, recurrent hitchhiking and background selection seem
unlikely to explain the observed patterns of diversity across species under our
understanding of the range of parameter estimates.

@@

First, I approximate the reduction in diversity using homogeneous background
selection and recurrent hitchhiking models
\citep{Kaplan1989-sc,Hudson1995-xc,Coop2012-cd}, when in reality, there is
genome-wide heterogeneity in functional density, recombination rates, and the
adaptive substitutions across species. Each of these factors mediate how
strongly linked selection impacts diversity across the genome. Despite these
model simplifications, the predicted reduction in diversity in \emph{Drosophila
melanogaster} is $85\%$ (when using $N_e$, not $N_c$), which is reasonably
close to the estimated $77\%$ from the more realistic model of Elyashiv et al.
that accounts for the actual position of substitutions, annotation features,
and recombination rate heterogeneity (though it should be noted that these both
use the same parameter estimates). Furthermore, even though my model fails to
capture the heterogeneity of functionality density and recombination rate in
real genomes, it is still extraordinary conservative, likely overestimating the
effects of linked selection to see if it could be capable of decoupling
diversity from census size and explain Lewontin's Paradox. This is in part
because the strong selection parameter estimates from \emph{Drosophila
melanogaster} used, but also because I assume that the effective population
size is equal to the census size.  Even then, this decoupling only occurs in
very high--census-size species, and implies that the diversity in the absence
of linked selection, $\pi_0$, is currently underestimated by several orders of
magnitude. Moreover, the study of \cite{Corbett-Detig2015-gt} did consider
recombination rate and functional density heterogeneity in estimating the
reduction due to linked selection across species, yet their predicted
reductions are orders of magnitude weaker than those considered here by
assuming that $N_e = N_c$ (Supplementary Figure \ref{figsupp:corbett_detig}B).
Overall, even with more realistic models of linked selection, current models of
linked selection seem fundamentally unable to fit the diversity--census-size
relationship.

@@

The data used in this study are collated from a variety of previously published
surveys. Of the 172 taxa with diversity estimates, 14 are from
\cite{Corbett-Detig2015-gt}, 96 are from \cite{Leffler2012-zj}, and 62
are from \cite{Romiguier2014-bp}. The Corbett-Detig et al. data is
estimated from four-fold degenerate sites, the Romiguier et al. data is
synonymous sites, and the Leffler et al. data is estimated predominantly from
silent, intronic, and non-coding sites. All types of diversity estimates from
\cite{Leffler2012-zj} were included to maximize the taxa in the study,
since the variability of diversity across functional categories is much less
than the diversity across taxa. Multiple diversity estimates per taxa were
averaged. The total recombination map length data were from both
\citeauthor{Stapley2017-fs} (2017; 127 taxa), and
\citeauthor{Corbett-Detig2015-gt} (2015; 9 taxa). Both studies used
sex-averaged recombination maps estimated with cross-based approaches; in some
cases errors in the original data were found, documented, and corrected. These
studies also included genome size estimates used to create Supplementary
Figures \ref{supfig:genome_size_popsize} and
\ref{supfig:genome_size_maplength}.

@@

Then, I re-estimated the relationship between body mass and population density
using the data in the appendix table of \cite{Damuth1987-sg}, which
includes 696 taxa with body mass and population density measurements across
mammals, fish, reptiles, amphibians, aquatic invertebrates, and terrestrial
arthropods. Though the abundance-body size relationship can be noisy at small
spatial or phylogenetic scales \citep[Chapter 5,][]{Gaston2008-pn}, across
deeply diverged taxa such as those included in this study and
\cite{Damuth1987-sg}, the relationship is linear and homoscedastic (see
Supplementary Figure \ref{figsupp:damuth}). Using Stan
\citep{Stan_Development_Team2020-ea}, I jointly estimated the relationship
between body mass from body length using the \cite{Romiguier2014-bp} taxa,
and used this relationship to predict body mass for the taxa in this study.
These body masses were then used to predict population density simultaneously,
using the \cite{Damuth1981-st} relationship. The code of this routine
(\texttt{pred\_popsize\_missing\_centered.stan}) is available in the GitHub
repository (\url{https://github.com/vsbuffalo/paradox_variation/}).

To estimate range, I first downloaded occurrence records from Global
Biodiversity Information Facility \citep{noauthor_2020-ey} using the
\texttt{rgbif} R package \citep{Chamberlain2014-up,Chamberlain2017-uz}.
Using the occurrence locations, I inferred whether a species was marine or
terrestrial, based on whether the majority of their recorded occurrences
overlapped a continent using \texttt{rnaturalearth} and the \texttt{sf}
packages \citep{South2017-db,Pebesma2018-bs}. For each taxon, I estimated
its range by finding the minimum $\alpha$-shape containing these occurrences.
The $\alpha$ parameters were set more permissive for marine species since
occurrence data for marine taxa were sparser. Then, I intersected the inferred
ranges for terrestrial taxa with continental polygons, so their ranges did not
overrun landmasses (and likewise with marine taxa and oceans). I inspected
diagnostic plots for each taxa for quality control (all of these plots are
available in \texttt{paradox\_variation} GitHub repository), and in some cases,
I manually adjusted the $\alpha$ parameter or manually corrected the range
based on known range maps (these changes are documented in the code
\texttt{data/species\_ranges.r} and \texttt{data/species\_range\_fixes.r}).
The range of \emph{C. elegans} was conservatively approximated as the area of
the Western US and Western Europe based on the map in \cite{Frezal2015-iw}.
\emph{Drosophila} species ranges are from the Drosophila Speciation Patterns
website, \citep{Yukilevich2012-vn,Yukilevich2017-jg}. To further validate
these range estimates, I have compared these to the qualitative range
descriptions \cite{Leffler2012-zj} (Supplementary Figure
\ref{figsupp:range-cats}) and compared my $\alpha$-shape method to a subset of
taxa with range estimates from IUCN Red List
(\cite{Chamberlain2020-tv,Iucn2020-ap}; Supplementary Figure
\ref{figsupp:redlist-ranges}). Each census population size
is then estimated as the product of range and density.

@@

First, I compared the relative carbon biomass in this study to the relative
carbon biomass on earth per phylum. This shows that this study's sample over
represents chordate biomass (by a factor of $\mathord{\sim}3$), and under
represents in arthropod biomass (by a factor of $0.02$) relative to the
proportion of carbon biomass of these phyla on earth (see column eight of
Supplementary Table \ref{supp:biomass-tbl}. Second, to check whether the carbon
biomass per phylum in the sample was broadly consistent with the total on earth
by phylum ($B_S$ for phylum $S$), I calculated the expected sample biomass if
species were sampled randomly from the total species in a phylum, ($B_S \times
\nicefrac{n_S}{T_S}$, where $n_S$ is the total number of species in the sample
in phylum $S$, $T_S$ is the total number of species in phylum $S$ on earth).
The fraction of total species on earth included in the sample in this study is
depicted in Supplementary Figure \ref{figsupp:species_counts}.

Next, I look at the ratio of sample biomass per phylum, $b_S$ to this expected
biomass per phylum (Supplementary Table \ref{supp:biomass-tbl}). The
consistency is quite close for this rough approach and the non-random sample of
taxa included in this study. The carbon biomass estimates for chordates implied
by the census size estimates are $\mathord{\sim}24$-fold higher than expected,
but is well within reasonable expectations given that the chordate sample
includes many larger-bodied domesticated species (and is a biased sample in
other ways).  Similarly, the implied arthropod carbon biomass is quite close to
what one would expect. Overall, these values indicate that the census size
estimates here do not lead to implied biomasses per phylum that are outside the
range of plausibility. For other population size consistency checks, see
Appendix \ref{app:popsize-valid}.

@@

Of the full dataset of 172 taxa with diversity and population size estimates, a
synthetic calibrated phylogeny was created for 166 species that appear in
phylogenies in DateLife project \citep{OMeara2020-ds,Sanchez-Reyes2019-io}.
This calibrated synthetic phylogeny was then subset for the analyses based on
what species had complete trait data. The diversity-population size
relationship assessed by a linear phylogenetic mixed-effects model implemented
in Stan \citep{Stan_Development_Team2020-ea}, according to the methods
described in (\cite{De_Villemereuil2014-kt}, see
\texttt{stan/phylo\_mm\_regression.stan} in the GitHub repository). This same
Stan model was used to estimate the same relationship between arthropod,
chordate, and mollusc subsets of the data, though a reduced model was used for
the chordate subset due to identifiability issues leading to poor MCMC
convergence (Supplementary Figure \ref{figsupp:div_popsize_post_phyla}).

The relationship between recombination map length and the logarithm of
population size is non-linear and heteroscedastic, and was fit using a
lognormal phylogenetic mixed-effects model on the 130 species with complete
data. Since social insects have longer recombination map lengths
\citep{Wilfert2007-dx}, social taxa were excluded when fitting this model.  All
\texttt{Rhat} \citep{Vehtari2019-po} values were below 1.01 and the effective
number of samples was over 1,000, consistent with good mixing; details about
the model are available in the GitHub repository
(\texttt{phylo\_mm\_lognormal.stan}). Continuous trait maps (\FIG{pcm}A and
Supplementary Information Figure \ref{figsupp:biconmap}) were created using
\texttt{phytools} \citep{Revell2012-zq}. Node-height tests were implemented
based on the methods in \texttt{Geiger} \citep{Pennell2014-nx,Harmon2008-tr},
and use robust regression to fit a linear relationship between phylogenetic
independent contrasts and branching times.

@@

\subsection{Citations}
\bibliography{biblio.bib}

>>>>

   \figsupp[The relationship between body mass and population density found by
   \cite{Damuth1987-sg}, which is used to predict population densities.]{The
     source of this data is appendix table of \cite{Damuth1987-sg}; the color
     indicates Damuth's original group labels. The dashed line was estimated
     using a lognormal regression model in Stan. References to each measurement
   are available in \cite{Damuth1987-sg}.}{\includegraphics[]{figures/damuth.pdf}}\label{figsupp:damuth}

@@

\figdata{The complete diversity and population size dataset for 172 metazoan taxa.}

@@

A major impediment in quantifying Lewontin's Paradox has been estimating census
population sizes across many taxa, especially for extremely abundant,
cosmopolitan species that define the upper limit of ranges. Previous work has
surveyed the literature for census size estimates
\citep{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range, body size, or
qualitative categories as proxies for census size
\citep{Corbett-Detig2015-gt,Leffler2012-zj}. To quantify the relationship
between genomic estimates of diversity and census population sizes, I first
approximate census population sizes for 172 metazoan taxa (\FIG{figure-nc}). My
approach predicts population densities from body sizes using a
previously-observed linear relationship that holds across metazoans (see
\FIGSUPP[figure-nc]{damuth}; \cite{Damuth1981-st,Damuth1987-sg}). Then, from
geographic occurrence data, I estimate range sizes.  Finally, I estimate
population size as the product of these predicted densities and range estimates
(see Methods and Materials: \nameref{sec:methods-popsize}). Note that the relationship between
population density and body size is driven by energy budgets, and thus reflects
macroecological equilibria \citep{Damuth1987-sg}. Consequently, population
sizes are underestimated for taxa like humans and their domesticated species,
and overestimated for species with anthropogenically reduced densities or
fragmented ranges. For example, the population size of \emph{Lynx lynx} is
likely around 50,000 \citep{Iucn2020-ap} which is around two orders of
magnitude smaller than my estimate. Additionally, the range size estimates do
not consider whether an area has unsuitable habitat, and thus may be
overestimated for species with particular niches or patchy habitats. While
these methods to estimate census size are crude and approximate, they can be
efficiently calculated for numerous taxa and are sufficient to estimate the
scale of Lewontin's Paradox (see \nameref{supp:popsize-valid} for more on
validation based on biomass and other approaches).

@@

\begin{align} 
  \pi_\text{BGS+HH} \approx \frac{\theta}{\nicefrac{1}{B(U, L)} + 2N_c S(\gamma, L, J)} 
  \label{eqn:div} 
\end{align}
%
where $\theta = 4N_c\mu$, $B(U,L)$ is the effect of background selection, and
$S(\gamma, L, J)$ is the rate of coalescence caused by sweeps (c.f.
\cite{Elyashiv2016-vt}, equation 1, \cite{Coop2012-cd} equation 20). Under
background selection models with recombination, the reduction is $B(U,L) =
\exp(\nicefrac{-U}{L})$ where $U$ is the per diploid genome per generation
deleterious mutation rate, and $L$ is the recombination map length
\citep{Hudson1994-oh,Hudson1995-xc,Nordborg1996-nq}. This BGS model is
similar to models of effective population size under polygenic fitness
variation, and can account for other modes of linked selection
(\cite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs}, see Appendix 
\ref{app:bgs}, \nameref{app:bgs}). The coalescent rate due to sweeps is $S(\gamma, L, J) =
\nicefrac{\gamma}{L} J$, where $\gamma$ is the number of adaptive substitutions
per generation, and $J$ is the probability a lineage is trapped by sweeps as
they occur across the genome (c.f. $J_{2,2}$ in equation 15 of
\cite{Coop2012-cd}).

@@

  \caption{(A) The observed relationship between recombination map length ($L$)
    and census size ($N_c$) across 136 species with complete data and known
    phylogeny. Triangle points indicate six social taxa excluded from the model
    fitting since these have adaptively higher recombination map lengths
    \citep{Wilfert2007-dx}. The dark gray line is the estimated
    relationship under a phylogenetic mixed-effects model, and the gray
    interval is the 95\% posterior average. (B) Points indicate the observed
    $\pi$--$N_c$ relationship across taxa shown in \FIG{pi-Nc},
    and the blue ribbon is the range of predicted diversity were $N_e = N_c$
    for $\mu = 10^{-8}$--$10^{-9}$,  and after accounting for the expected
  reduction in diversity due to background selection and recurrent hitchhiking
under \emph{Drosophila melanogaster} parameters. In both plots, point color
indicates phylum.}\label{fig:linksel}

@@

Previous work has found that the impact of linked selection increases with
$N_c$ (\cite{Corbett-Detig2015-gt}; see \FIGSUPP[linksel]{reductions}A), and it
is often thought that this is driven by higher rates of adaptive substitutions
in larger populations \citep{Ohta1992-yi}, despite equivocal evidence
\citep{Galtier2016-dq}.  However, there is another mechanism by which species
with larger population sizes might experience a greater impact of linked
selection: recombinational map length, $L$, is known to correlate with body
mass \citep{Burt1987-tq} and thus varies inversely with population size. As
this is a critical parameter that determines the genome-wide impact of both
hitchhiking and background selection, I examine the relationship between
recombination map length ($L$) and census population size ($N_c$) across taxa,
using available estimates of map lengths across species
\citep{Stapley2017-fs,Corbett-Detig2015-gt}. I find a significant non-linear
relationship using phylogenetic mixed-effects models ( \FIG{linksel}A; see
Methods and Materials: \nameref{sec:methods-pcm}).  There is also a correlation
between map length and genome size (\FIGSUPP[linksel]{genome-size-maplen}) and
genome size and population size (\FIGSUPP[linksel]{genome-size-Nc}). These
findings are consistent with both the hypothesis that non-adaptive processes
increase genome size in small-$N_e$ species \citep{Lynch2003-bn} which in turn
could increase map lengths, as well as the hypothesis that map lengths are
adaptively longer to more efficiently select against deleterious alleles
\citep{Roze2021-ya}.  Overall, the negative relationship between map length and
census size indicates linked selection is expected to be stronger in short map
length, high-$N_c$ species.

Then, I predict the expected diversity ($\pi_{BGS+HH}$) under background
selection and hitchhiking, were $N_e= N_c$, and assuming all species had the
rate of sweeps and strength of BGS as \emph{D. melanogaster}. Since neutral
mutation rates $\mu$ are unknown and vary across species, I calculate the range
of predicted $\pi_{BGS+HH}$ estimates for $\mu = 10^{-8}$--$10^{-9}$ (using the
four-alleles model, \cite{Tajima1996-rb}), and compare this to the observed
relationship between $\pi$ and $N_c$ in \FIG{linksel}B.  Under these
parameters, linked selection begins to appreciably depress diversity around
$N_c \approx 10^9$, since $S \approx 10^{-8}$--$10^{-9}$ and linked selection
dominates drift when $S > \nicefrac{1}{2N}$.  Overall, this reveals two
problems for the hypothesis that linked selection could solve Lewontin's
Paradox. First, low to mid-$N_c$ species (census sizes between
$10^{4}$--$10^{10}$) have sufficiently long map lengths that their diversity
levels are only moderately reduced by linked selection, leading to a wide gap
between predicted and observed diversity levels. For this not to be the case,
the parameters that determine the strength of background selection and
recurrent hitchhiking would need to be \emph{higher} among these species than
in \emph{Drosophila melanogaster}. This would require that the rate of adaptive
mutations or the deleterious mutation rate be orders of magnitude higher for
species within this range than in \emph{Drosophila}, which is incompatible with
the rate of adaptive substitutions across species \citep{Galtier2016-dq} and
mutation rates \citep{Lynch2010-ki}. Furthermore, linked selection has been
quantified in humans, which fall in this census size range, and has been found
to be relatively weak
\citep{McVicker2009-ax,Hernandez2011-gs,Hellmann2008-ic,Cai2009-by,Boyko2008-tj}.
Second, while hitchhiking and BGS can reduce predicted diversity levels for
high-$N_c$ species ($N_c > 10^{10}$) to observed levels, this would imply
available estimates of $\pi_0$ are underestimated by several orders of
magnitude in \emph{Drosophila} (\FIGSUPP[linksel]{reductions}B). The high
reductions in $\pi$ predicted here (compared to those of
\cite{Elyashiv2016-vt}) are a result of using $N_c$, rather than $N_e =
\nicefrac{\pi_0}{4 \mu}$ in the denominator of Equation \eqref{eqn:div}, which
leads to a very high rate of sweeps in the population.  I do not consider
selective interference, though the saturation of adaptive substitutions per
Morgan would only act to limit the reduction in diversity
\citep{Weissman2012-aa}, and thus these results are conservative.  Overall,
while linked selection could decouple diversity from census size for high-$N_c$
species, recurrent hitchhiking and background selection seem unlikely to
explain the observed patterns of diversity across species under our
understanding of the range of parameter estimates.

@@

First, I approximate the reduction in diversity using homogeneous background
selection and recurrent hitchhiking models
\citep{Kaplan1989-sc,Hudson1995-xc,Coop2012-cd}, when in reality, there is
genome-wide heterogeneity in functional density, recombination rates, and the
adaptive substitutions across species. Each of these factors mediate how
strongly linked selection impacts diversity across the genome. Despite these
model simplifications, the predicted reduction in diversity in \emph{Drosophila
melanogaster} is $85\%$ (when using $N_e$, not $N_c$), which is reasonably
close to the estimated $77\%$ from the more realistic model of Elyashiv et al.
that accounts for the actual position of substitutions, annotation features,
and recombination rate heterogeneity (though it should be noted that these both
use the same parameter estimates). Furthermore, even though my model fails to
capture the heterogeneity of functionality density and recombination rate in
real genomes, it is still extraordinary conservative, likely overestimating the
effects of linked selection to see if it could be capable of decoupling
diversity from census size and explain Lewontin's Paradox. This is in part
because the strong selection parameter estimates from \emph{Drosophila
melanogaster} used, but also because I assume that the effective population
size is equal to the census size.  Even then, this decoupling only occurs in
very high--census-size species, and implies that the diversity in the absence
of linked selection, $\pi_0$, is currently underestimated by several orders of
magnitude. Moreover, the study of \cite{Corbett-Detig2015-gt} did consider
recombination rate and functional density heterogeneity in estimating the
reduction due to linked selection across species, yet their predicted
reductions are orders of magnitude weaker than those considered here by
assuming that $N_e = N_c$ (\FIGSUPP[linksel]{reductions}B).
Overall, even with more realistic models of linked selection, current models of
linked selection seem fundamentally unable to fit the diversity--census-size
relationship.

@@

The data used in this study are collated from a variety of previously published
surveys. Of the 172 taxa with diversity estimates, 14 are from
\cite{Corbett-Detig2015-gt}, 96 are from \cite{Leffler2012-zj}, and 62 are from
\cite{Romiguier2014-bp}. The Corbett-Detig et al. data is estimated from
four-fold degenerate sites, the Romiguier et al. data is synonymous sites, and
the Leffler et al. data is estimated predominantly from silent, intronic, and
non-coding sites. All types of diversity estimates from \cite{Leffler2012-zj}
were included to maximize the taxa in the study, since the variability of
diversity across functional categories is much less than the diversity across
taxa. Multiple diversity estimates per taxa were averaged. The total
recombination map length data were from both \citeauthor{Stapley2017-fs} (2017;
127 taxa), and \citeauthor{Corbett-Detig2015-gt} (2015; 9 taxa). Both studies
used sex-averaged recombination maps estimated with cross-based approaches; in
some cases errors in the original data were found, documented, and corrected.
These studies also included genome size estimates used to create
\FIGSUPP[linksel]{genome-size-maplen} and \FIGSUPP[linksel]{genome-size-Nc}.

@@

Then, I re-estimated the relationship between body mass and population density
using the data in the appendix table of \cite{Damuth1987-sg}, which
includes 696 taxa with body mass and population density measurements across
mammals, fish, reptiles, amphibians, aquatic invertebrates, and terrestrial
arthropods. Though the abundance-body size relationship can be noisy at small
spatial or phylogenetic scales \citep[Chapter 5,][]{Gaston2008-pn}, across
deeply diverged taxa such as those included in this study and
\cite{Damuth1987-sg}, the relationship is linear and homoscedastic (see
\FIGSUPP[figure-nc]{damuth}). Using Stan
\citep{Stan_Development_Team2020-ea}, I jointly estimated the relationship
between body mass from body length using the \cite{Romiguier2014-bp} taxa,
and used this relationship to predict body mass for the taxa in this study.
These body masses were then used to predict population density simultaneously,
using the \cite{Damuth1981-st} relationship. The code of this routine
(\texttt{pred\_popsize\_missing\_centered.stan}) is available in the GitHub
repository (\url{https://github.com/vsbuffalo/paradox_variation/}).

To estimate range, I first downloaded occurrence records from Global
Biodiversity Information Facility \citep{noauthor_2020-ey} using the
\texttt{rgbif} R package \citep{Chamberlain2014-up,Chamberlain2017-uz}.  Using
the occurrence locations, I inferred whether a species was marine or
terrestrial, based on whether the majority of their recorded occurrences
overlapped a continent using \texttt{rnaturalearth} and the \texttt{sf}
packages \citep{South2017-db,Pebesma2018-bs}. For each taxon, I estimated its
range by finding the minimum $\alpha$-shape containing these occurrences.  The
$\alpha$ parameters were set more permissive for marine species since
occurrence data for marine taxa were sparser. Then, I intersected the inferred
ranges for terrestrial taxa with continental polygons, so their ranges did not
overrun landmasses (and likewise with marine taxa and oceans). I inspected
diagnostic plots for each taxa for quality control (all of these plots are
available in \texttt{paradox\_variation} GitHub repository), and in some cases,
I manually adjusted the $\alpha$ parameter or manually corrected the range
based on known range maps (these changes are documented in the code
\texttt{data/species\_ranges.r} and \texttt{data/species\_range\_fixes.r}).
The range of \emph{C. elegans} was conservatively approximated as the area of
the Western US and Western Europe based on the map in \cite{Frezal2015-iw}.
\emph{Drosophila} species ranges are from the Drosophila Speciation Patterns
website, \citep{Yukilevich2012-vn,Yukilevich2017-jg}. To further validate these
range estimates, I have compared these to the qualitative range descriptions
\cite{Leffler2012-zj} (\FIGSUPP[figure-nc]{range-cats}) and compared my
$\alpha$-shape method to a subset of taxa with range estimates from IUCN Red
List (\cite{Chamberlain2020-tv,Iucn2020-ap};
\FIGSUPP[figure-nc]{redlist-ranges}). Each census population size is then
estimated as the product of range and density.

@@

First, I compared the relative carbon biomass in this study to the relative
carbon biomass on earth per phylum. This shows that this study's sample over
represents chordate biomass (by a factor of $\mathord{\sim}3$), and under
represents in arthropod biomass (by a factor of $0.02$) relative to the
proportion of carbon biomass of these phyla on earth (see column eight of
Table \ref{supp:biomass-tbl}. Second, to check whether the carbon
biomass per phylum in the sample was broadly consistent with the total on earth
by phylum ($B_S$ for phylum $S$), I calculated the expected sample biomass if
species were sampled randomly from the total species in a phylum, ($B_S \times
\nicefrac{n_S}{T_S}$, where $n_S$ is the total number of species in the sample
in phylum $S$, $T_S$ is the total number of species in phylum $S$ on earth).
The fraction of total species on earth included in the sample in this study is
depicted in \FIGSUPP[figure-nc]{species-counts}.

Next, I look at the ratio of sample biomass per phylum, $b_S$ to this expected
biomass per phylum (Table \ref{supp:biomass-tbl}). The
consistency is quite close for this rough approach and the non-random sample of
taxa included in this study. The carbon biomass estimates for chordates implied
by the census size estimates are $\mathord{\sim}24$-fold higher than expected,
but is well within reasonable expectations given that the chordate sample
includes many larger-bodied domesticated species (and is a biased sample in
other ways).  Similarly, the implied arthropod carbon biomass is quite close to
what one would expect. Overall, these values indicate that the census size
estimates here do not lead to implied biomasses per phylum that are outside the
range of plausibility. For other population size consistency checks, see
Appendix \ref{app:popsize-valid}.

@@

Of the full dataset of 172 taxa with diversity and population size estimates, a
synthetic calibrated phylogeny was created for 166 species that appear in
phylogenies in DateLife project \citep{OMeara2020-ds,Sanchez-Reyes2019-io}.
This calibrated synthetic phylogeny was then subset for the analyses based on
what species had complete trait data. The diversity-population size
relationship assessed by a linear phylogenetic mixed-effects model implemented
in Stan \citep{Stan_Development_Team2020-ea}, according to the methods
described in (\cite{De_Villemereuil2014-kt}, see
\texttt{stan/phylo\_mm\_regression.stan} in the GitHub repository). This same
Stan model was used to estimate the same relationship between arthropod,
chordate, and mollusc subsets of the data, though a reduced model was used for
the chordate subset due to identifiability issues leading to poor MCMC
convergence (\FIGSUPP[pcm]{div-popsize-post-phyla}).

The relationship between recombination map length and the logarithm of
population size is non-linear and heteroscedastic, and was fit using a
lognormal phylogenetic mixed-effects model on the 130 species with complete
data. Since social insects have longer recombination map lengths
\citep{Wilfert2007-dx}, social taxa were excluded when fitting this model.  All
\texttt{Rhat} \citep{Vehtari2019-po} values were below 1.01 and the effective
number of samples was over 1,000, consistent with good mixing; details about
the model are available in the GitHub repository
(\texttt{phylo\_mm\_lognormal.stan}). Continuous trait maps (\FIG{pcm}A,
\FIGSUPP[pcm]{anc-maplen}, and \FIGSUPP[pcm]{anc-species}) were created using
\texttt{phytools} \citep{Revell2012-zq}. Node-height tests were implemented
based on the methods in \texttt{Geiger} \citep{Pennell2014-nx,Harmon2008-tr},
and use robust regression to fit a linear relationship between phylogenetic
independent contrasts and branching times.

@@

\bibliography{biblio.bib}
